Ground state energy of quantum dots using the coupled cluster method

### 2018-06-08 project_2/article/article.tex
Thorough work-through.

\begin{abstract}
    Something about coupled-cluster... Preferably doubles.
\end{abstract}

@@

\section{Introduction}
    %TODO: Add explanation on why the many-body problem becomes intractable.
    % This can be shown by demonstrating that the true wavefunction of the system
    % must be a linear combination of all possible Slater detemerminants built up
    % from permutations of the single particle functions in the basis. Thus the
    % energy equation becomes too large.
    In this project we will study the ground state energy of quantum dots.

@@

    \subsection{Quantum Dots}

        Some consideration should be given towards why the system described above
        is a quantum dot, and why we should care about quantum dots.
        There are at least two things we refer to as quantum dots. The
        first one is a man-made device, also known as artificial atoms,
        and the other is the simplest theoretical construct of such a device.

        Aritifical atom quantum dots are semiconducting crystals, typically GaAs
        relatively easy to build in a laboratory. Here, some free electrons are 
        confined to a volume, either by a physical barrier like an insulator
        or an electromagnetic field\cite{reimann2002electronic}. Due to the
        confinement of the electrons, the energy levels of a quantum dot becomes
        quantized, and they therefore hold many properties similar to naturally
        occuring quantized systems, which is the reason why quantum dots have
        attracted much interest.

        The theoretical approximation of a quantum dot is impossible to make perfectly
        accurate, yet simple. There are some important effects that should be
        included such the potential arising as a result of the confinement, and
        the interaction between particles. We therefore model our system with an
        electron-electron interaction, given by the two-body operator in
        \autoref{eq:eq:two_body_operator}, and with we approximate the confining
        potential by the harmonic oscillator potential, given by the second term
        in the one-body Hamiltonian in \autoref{eq:eq:one_body_hamiltonian}. The
        last term in the one-body Hamiltonian is the kinetic energy of particle
        in question, essential in any system.

    \subsection{Second quantization}
        Employing the creation operators, $\acr{p}$, and the destruction
        operators, $\ade{p}$, we can write the Hamiltonian as
        \begin{align}
            H
            &=
            \sum_{pq}h_{q}^{p}\acr{p}\ade{q}
            + \frac{1}{4}\sum_{pqrs}\bra{pq}\ket{rs}\acr{p}\acr{q}\ade{s}\ade{r},
        \end{align}
        where the sums are general indices over all $L$ basis states and the
        matrix elements are defined as
        \begin{gather}
            h^{p}_{q} \equiv \bra{p}h\ket{q}, \\
            \bra{pq}\ket{rs} \equiv \bra{pq}u\ket{rs} - \bra{pq}u\ket{sr}.
        \end{gather}
        Note that we use the chemists notation to label the antisymmetric matrix
        elements.

    \subsection{The coupled cluster approximation}
        We approximate the true wavefunction, $\ket{\Psi}$, of the system by the
        coupled cluster wavefunction, $\ket{\Psi_{\text{CC}}}$, defined by
        \begin{align}
            \ket{\Psi_{\text{CC}}}
            &\equiv e^{T}\ket{\Phi_0}
            = \para{
                \sum_{i = 0}^n
                \frac{1}{n!}T^n
            }\ket{\Phi_0},
        \end{align}
        where the \emph{cluster operator}, $T$, is given by a sum of
        $p$-excitation operators labeled $T_p$. They consist of \emph{cluster
        amplitudes}, $t_{i\dots}^{a\dots}$, and creation and annihilation
        operators.
        \begin{align}
            T &= T_1 + T_2 + \dots + T_p \\
            &=
            \sum_{ia}t_i^a\acr{a}\ade{i}
            + \para{\frac{1}{2!}}^2\sum_{ijab}
            t_{ij}^{ab}\acr{a}\acr{b}\ade{i}\ade{j}
            + \dots.
        \end{align}
        In the doubles approximation we limit the cluster operator to
        \begin{align}
            T \equiv T_2
            = \frac{1}{4}\sum_{ijab}t_{ij}^{ab}\acr{a}\acr{b}\ade{j}\ade{i}.
            \label{eq:T_2}
        \end{align}
        The first part of the coupled cluster method consists of constructing
        the cluster amplitudes using the \emph{amplitude equations}. After we
        have found the amplitudes we can compute the energy.

@@

    \subsection{Coupled cluster amplitude equations}
        In order for us to solve the energy equation using the coupled cluster
        approximation we need to figure out what the cluster amplitudes,
        $t_{ij\dots}^{ab\dots}$, are. This is done by projecting
        \autoref{eq:clean_schrodinger} onto an excited Slater determinant, i.e.,
        \begin{align}
            \bra{\Phi_{ij\dots}^{ab\dots}}e^{-T}He^{T}\kslat
            = 0.
        \end{align}
        Note that in the amplitude equations we can use both the regular and the
        normal ordered Hamiltonian. They are equal as the reference energy term
        disappears due to \autoref{eq:excited_overlap}. The order of the
        excitation in the projection determines the order of the amplitudes you
        will find. In our case we are only interested in the second order
        ampltiudes found in the doubles approximation, hence we will solve the
        equation
        \begin{align}
            \bra{\Phi_{ij}^{ab}}e^{-T}H_N e^{T}\kslat = 0,
            \label{eq:doubles_amplitude_1}
        \end{align}
        to find an expression that can be used to solve for $t_{ij}^{ab}$.
        By employing the Baker-Campbell-Haussdorf (BCH) expansion,
        while setting $T= T_2$, we find
        \begin{equation}
                \bar{H}
                    = \Big(
                            H_N + H_NT_2
                            + \frac{1}{2}H_NT_2^2 \Big)_C.
        \label{eq:normal_order_expansion}
        \end{equation}
        The subscript $c$ indicates that only those terms in which the
        Hamiltonian is connected\footnote{In a Wick's theorem sense}
        to every cluster operator on its right should be included. Since
        the Hamiltonian contains at most four annihilation and creation
        operators, $H_N$ can connect to at most four cluster operators
        at once. Therefor, the BCH expansion must truncate at the
        fourth-order terms.

@@

        The we can define a new function consisting of all of \autoref{eq:CCD_amp_1}
        except term number two and three,
        \begin{equation}
        \begin{aligned}
            g(u, t) =&
                    u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab) - f^k_jt^{ab}_{ik}P(ij) \\
                 +& \frac{1}{4}t^{cd}_{ij} t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}t^{cd}_{ij} u^{ab}_{cd} \\
                 +& \frac{1}{2}t^{cd}_{jm} t^{ab}_{in} u^{mn}_{cd} P(ij) - \frac{1}{2}t^{ac}_{nm} t^{bd}_{ij} u^{nm}_{cd} P(ab) \\
                 +& t^{ac}_{im} t^{bd}_{jn} u^{mn}_{cd} P(ij) + t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
                 -& \frac{1}{2}t^{ab}_{im} u^{mn}_{jn}.
        \end{aligned}
        \end{equation}

@@

        \subsubsection{Intermediate Computations}
        Looking closely at the amplitude equation in (\ref{eq:CCD_amp_1}) one might come
        to realize that it contains that this equation contains many of the same structures in
        several of the terms. This warrants the search for an algebraic transformation of the
        CCD amplitude equation that has the potential to reduce the amount of floating point
        operations needed to compute it. As it turns out, such terms exist and they will decrease
        the computing time necessary by an order of magnitude. We will define the following
        "intermediates",
        \begin{align}
            \label{eq:intermediate1}
            \chi^{ab}_{cd} &= \frac{1}{4}t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}u^{ab}_{cd} \\
            \label{eq:intermediate2}
            \chi^n_j &= \frac{1}{2}t^{cd}_{jm} u^{mn}_{cd} \\
            \label{eq:intermediate3}
            \chi^a_d &= \frac{1}{2} t^{ac}_{nm} u^{nm}_{cd} \\
            \label{eq:intermediate4}
            \chi^{bm}_{jc} &= u^{bm}_{jc} + \frac{1}{2}t^{bd}_{jn}u^{mn}_{cd}
        \end{align}
        These intermediate structures will allow us to rewrite \autoref{eq:CCD_amp_1} to,
        \begin{equation}
        \begin{aligned}
            0 =& u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab) - f^k_jt^{ab}_{ik}P(ij) \\
              &+ t^{cd}_{ij}\chi^{ab}_{cd} + t^{ab}_{in}\chi^n_jP(ij)
               - t^{bd}_{ij}\chi^a_dP(ab) \\
              &+ t^{ac}_{im}\chi^{bm}_{jc}P(ab)P(ij) + \frac{1}{2}t^{ab}_{im}u^{mn}_{jn}.
        \end{aligned}
        \end{equation}
        The importance of this "trick" will become apparent in due time.

@@

        \subsubsection{Constructing the Hartree-Fock basis}
            Having found the matrix elements of $h$ and $u$ we can now use the
            \emph{self-consistent field iteration} method to construct $h$ and
            $u$ in a \emph{restricted Hartree-Fock} basis. This will yield a
            better estimate to the actual, unknown basis functions of the
            system.

            % See Step 3 in:
            % http://sirius.chem.vt.edu/wiki/doku.php?id=crawdad:programming:project4

            The neatest, yet arguably the most abstract, way to write the Hartree-Fock
            equations for electron $i$\footnote{This is weird, as electron
            should be indistinguishable and therefore impossible to label.} is
            \begin{equation}
                f_i \varphi_i = \varepsilon_i \varphi_i,
            \end{equation}
            where $f_i$ is the Fock operator, $\varphi_i$ are eigenstates of the Fock
            operator consiting of a set of one-electorn wave funcitons, called the
            Hartree-Fock molecular orbitals, and $\varepsilon_i$ are the eigenenergies
            of the Fock operator. The fock operator, in matrix notation, is given by
            \begin{equation}
                F_{pa} = H_{pq} +  J(D)_{pq} - \frac{1}{2}K(D)_{pq},
            \end{equation}
            where $h_i$ is the one-body operatore, while the two-body operator is
            divided into what we call a direct part,
            \begin{equation}
                J(D)_{pq} = \braket{pq}{rs}D_{sr},
            \end{equation}
            and an exchange part,
            \begin{equation}
                K(D)_{pq} = \braket{ps}{rq}D_{sr}.
            \end{equation}
            The direct part of the two-body operator is comparable to classical Coloumb
            repulsion, while the exchange does not have a classical analog as it arises
            from the antisymmetry requirement of the wavefunction. $D_{sr}$ is the
            density matrix of the system.

@@

            Because the Hartree-Fock method is a variational method, the initial
            wavefunction made up of the single-partrice wavefunctions $\varphi_i$
            is varied until the minimal energy is found. The transformations necessary
            to make all these variations will now be contained in $C$. So if we start from
            a Harmonic Oscillator (HO) basis we can transform into a better basis, called the
            Hartree-Fock (HF) basis,
            \begin{equation}
                \ket{\Psi}^{\text{HF}}_p = \sum_{k=1}^L C_{kp} \ket{\Psi}_k^{\text{HO}}.
            \end{equation}
            We will see how the coupled cluster method with double excitations works with
            both the HO basis and the HF basis.

@@

    Because we have employed a mixture of Python and C++ with Cython interfaces,
    the code sturcture may be difficult to undertand for someone without Cython
    experience, but we have neverteless tried to make the usage of the software as
    painless as possible.

@@

        The software is easy to install by first cloning the GitHub reopsitory,
        \begin{bash}
            git clone git@github.com:Schoyen/FYS4411.git
        \end{bash}
        Then all you need to do is change directory to the project code folder
        where a \lstinline{MakeFile} is included so you only need to build and
        install,
        \begin{bash}
            cd FYS4411/project_2/coupled-cluster
            make build
            make install
        \end{bash}
        Now everything should be able to run from anywhere on your computer.
        To ensure that all requirements of the program are satified you can
        install with \lstinline{make installr} instead.

        We have included a sample python script in \autoref{app:sample_script}
        that uses almost all methods and classes in the package.

@@

        Both in \mintinline{python}{CoupledClusterDoublesSparse} and
        \mintinline{python}{CoupledClusterDoublesOptimized} we make use
        of the intermediates in equations \ref{eq:intermediate1} to
        \ref{eq:intermediate4}. In the optimized class, special care was
        taken to compute

@@

        % TODO: Add short introduction to each table and link to them. Leave the
        % discussion of the results in the discussion section.

@@

        \begin{table}
            \centering
            \caption{Here we look at $N = 12$ particles ($R = 3$ shells).  We did
            not achieve convergence using the harmonic oscillator basis for the
            lower frequency values and large number of shells. No convergence is
            marked as {\nan} in the table.}
            \begin{ruledtabular}
                \begin{tabular}{c|c|ccc}
                    $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF) \\
                    \hline
                          & $3$ & $46.361130$ & $46.361130$ & $46.361130$ \\
                          & $4$ & $43.663267$ & $45.837079$ & $43.309845$ \\
                          & $5$ & $41.108851$ & $45.456883$ & $40.654710$ \\
                          & $6$ & $40.750512$ & \nan & $40.068340$ \\
                    $0.5$ & $7$ & $40.302719$ & \nan & $39.508500$ \\
                          & $8$ & $40.263752$ & \nan & $39.399128$ \\
                          & $9$ & $40.216688$ & \nan & $39.329311$ \\
                          & $10$ & $40.216252$ & \nan & $39.309409$ \\
                          & $11$ & $40.216195$ & \nan & $39.296007$ \\
                          & $12$ & $40.216165$ & \nan & $39.285968$ \\
                    \hline
                          & $3$ & $73.765549$ & $73.765549$ & $73.765549$ \\
                          & $4$ & $70.673849$ & $73.314476$ & $70.324250$ \\
                          & $5$ & $67.569930$ & $72.990679$ & $67.031096$ \\
                          & $6$ & $67.296869$ & \nan & $66.526677$ \\
                    $1.0$ & $7$ & $66.934745$ & \nan & $66.049564$ \\
                          & $8$ & $66.923094$ & \nan & $65.972157$ \\
                          & $9$ & $66.912244$ & \nan & $65.921205$ \\
                          & $10$ & $66.912035$ & \nan & $65.889281$ \\
                          & $11$ & $66.911365$ & \nan & $65.866715$ \\
                          & $12$ & $66.911364$ & \nan & $65.849776$ \\
                    \hline
                          & $3$ & $120.722260$ & $120.722260$ & $120.722260$ \\
                          & $4$ & $117.339642$ & $120.296556$ & $116.995036$ \\
                          & $5$ & $113.660396$ & $120.007146$ & $113.048934$ \\
                          & $6$ & $113.484866$ & $119.759037$ & $112.658821$ \\
                    $2.0$ & $7$ & $113.247601$ & $119.662199$ & $112.309482$ \\
                          & $8$ & $113.246579$ & $119.584733$ & $112.235521$ \\
                          & $9$ & $113.246303$ & $119.524394$ & $112.181828$ \\
                          & $10$ & $113.245854$ & $119.472283$ & $112.140661$ \\
                          & $11$ & $113.245256$ & $119.430353$ & $112.109973$ \\
                          & $12$ & $113.245183$ & $119.394712$ & $112.085683$ \\
                    \hline
                    $5.0$ & $3$ & $242.334879$ & $242.334879$ & $242.334879$ \\
                          & $4$ & $238.739591$ & $241.927593$ & $238.394598$ \\
                          & $5$ & $234.352741$ & $241.663950$ & $233.680649$ \\
                          & $6$ & $234.282331$ & $241.507595$ & $233.425243$ \\
                          & $7$ & $234.194820$ & $241.390525$ & $233.226529$ \\
                          & $8$ & $234.194059$ & $241.293417$ & $233.137933$ \\
                          & $9$ & $234.190797$ & $241.221056$ & $233.070205$ \\
                          & $10$ & $234.190714$ & $241.158896$ & $233.020009$ \\
                          & $11$ & $234.190665$ & $241.109926$ & $232.980634$ \\
                          & $12$ & $234.190553$ & $241.068091$ & $232.948528$
                \end{tabular}
            \end{ruledtabular}
            \label{tab:N12}
        \end{table}

>>>>

\begin{abstract}
    The coupled cluster method for estimating the ground state energy of a many-body
    system is arguably the most important post-Hartree-Fock method. We have implemented
    the coupled cluster method with double excitations (CCD). We find that it does not
    perform well with a simple harmonic oscillator basis, but the results vastly improve 
    when it is provided a basis from a restricted Hatree-Fock self-consistent field method
    run as a preliminary computation. For a system consisting of $N=2$ interacting
    particles, using 12 orbital shells we get a ground stated energy of $3.005970$, very
    close to the analytical ``true answer'' of $3$\cite{taut1994two}. 
\end{abstract}

@@

\section{Introduction}
    In this project we will compute the ground state energy of quantum dots, a
    construct that has become a popular area of study in recent decades, both from
    a theoretical and an experimental point of view. We achieve the ground state
    computation by implementing a coupled cluster amplitude equation solver
    which is used in succession with a restricted Hartree-Fock self-consistent
    field method.

    Regardless of how sophisticated the method used, the many-body problem remains
    quite intractable. The true wavefunction of any system will be a linear 
    combination of all possible slater-type orbitals one can model. This means that
    in order to represent a system correctly, one would need a basis set of infinite
    size. This is an obvious problem when we solve problems numerically, because we
    obviously need to limit the size of the basis we use because we have limitations
    in memeory and computating power. Luckily, great minds have bestowed a wide 
    selection of very god numerical methods upon the world, which the coupled cluster
    is a part of.

    First, we provide a thorough theoretical outline of how we model a quantum dot, 
    the mathematical formalism behind many-body problems and outline how the 
    computational methods we employ both turn into iterative schemes that are 
    easy to represent numerically. Second, we provide a brief overview of the
    program we have constructed which is in the form an installable python package.
    Third, we have employed the methods on a large parameter space presenting 
    results for systems consisting of $N= \{2, 6, 12, 20\}$ particles. Fourth,
    we discuss the most interesting aspects of the results and try to shed some
    light over the shortcomings as well as the beneficial features of the different
    metods. Lastly, we end on some summary remarks.

@@

    \subsection{\emph{Motivational Interlude}: Quantum Dots}

        The system described above is a theoretical approximation of a
        quantum dot. Some consideration should be given to an explanation of 
        why the system described above is a quantum dot, and why we should
        care about quantum dots.
        There are at least two things we refer to as quantum dots. The
        first one is a man-made device, also known as artificial atoms,
        and the other is the simplest theoretical construct of such a device.

        Aritifical atom quantum dots are semiconducting crystals, typically GaAs,
        and they are
        relatively easy to build in a laboratory. Some free electrons are 
        confined to a volume, either by a physical barrier, like an insulator,
        or an electromagnetic field\cite{reimann2002electronic}. Due to the
        confinement of the electrons, the energy levels of a quantum dot becomes
        quantized, and they therefore hold many properties similar to naturally
        occuring quantized systems, which is the reason why quantum dots have
        attracted much interest.

        The theoretical approximation of a quantum dot is impossible to make perfectly
        accurate, yet simple. There are some important effects that should be
        included such the potential arising as a result of the confinement, and
        the interaction between particles. We therefore model our system with an
        electron-electron interaction, given by the two-body operator in
        \autoref{eq:two_body_operator}, and with we approximate the confining
        potential by the harmonic oscillator potential, given by the second term
        in the one-body Hamiltonian in \autoref{eq:one_body_hamiltonian}. The
        last term in the one-body Hamiltonian is the kinetic energy of particle
        in question, essential in any system.

    \subsection{Second quantization}
        Employing the creation operators, $\acr{p}$, and the destruction
        operators, $\ade{p}$, we can write the Hamiltonian as
        \begin{align}
            H
            &=
            \sum_{pq}h_{q}^{p}\acr{p}\ade{q}
            + \frac{1}{4}\sum_{pqrs}\bra{pq}\ket{rs}\acr{p}\acr{q}\ade{s}\ade{r},
        \end{align}
        where the sums are general indices over all $L$ basis states and the
        matrix elements are defined as
        \begin{gather}
            h^{p}_{q} \equiv \bra{p}h\ket{q}, \\
            \bra{pq}\ket{rs} \equiv \bra{pq}u\ket{rs} - \bra{pq}u\ket{sr}.
        \end{gather}
        Note that we use the ``chemists notation'' to label the antisymmetric matrix
        elements.

    \subsection{The coupled cluster approximation}
        We approximate the true wavefunction, $\ket{\Psi}$, of the system by the
        coupled cluster wavefunction, $\ket{\Psi_{\text{CC}}}$, defined by
        \begin{align}
            \ket{\Psi_{\text{CC}}}
            &\equiv e^{T}\ket{\Phi_0}
            = \para{
                \sum_{i = 0}^n
                \frac{1}{n!}T^n
            }\ket{\Phi_0},
        \end{align}
        where the \emph{cluster operator}, $T$, is given by a sum of
        $p$-excitation operators labeled $T_p$. They consist of \emph{cluster
        amplitudes}, $t_{i\dots}^{a\dots}$, and creation and annihilation
        operators.
        \begin{align}
            T &= T_1 + T_2 + \dots + T_p \\
            &=
            \sum_{ia}t_i^a\acr{a}\ade{i}
            + \para{\frac{1}{2!}}^2\sum_{ijab}
            t_{ij}^{ab}\acr{a}\acr{b}\ade{i}\ade{j}
            + \dots.
        \end{align}
        In our approximation we limit the cluster operator to only include
        double excitations,
        \begin{align}
            T \equiv T_2
            = \frac{1}{4}\sum_{ijab}t_{ij}^{ab}\acr{a}\acr{b}\ade{j}\ade{i}.
            \label{eq:T_2}
        \end{align}
        The first part of the coupled cluster method consists of constructing
        the cluster amplitudes using the \emph{amplitude equations}. After we
        have found the amplitudes we can compute the energy.

@@

    \subsection{Coupled cluster amplitude equations}
        In order for us to solve the energy equation using the coupled cluster
        approximation we need to figure out what the cluster amplitudes,
        $t_{ij\dots}^{ab\dots}$, are. This is done by projecting
        \autoref{eq:clean_schrodinger} onto an excited Slater determinant, i.e.,
        \begin{align}
            \bra{\Phi_{ij\dots}^{ab\dots}}e^{-T}He^{T}\kslat
            = 0.
        \end{align}
        Note that in the amplitude equations we can use both the regular and the
        normal ordered Hamiltonian. They are equal as the reference energy term
        disappears due to \autoref{eq:excited_overlap}. The order of the
        excitation in the projection determines the order of the amplitudes you
        will find. In our case we are only interested in the second order
        ampltiudes found in the doubles approximation, hence we will solve the
        equation
        \begin{align}
            \bra{\Phi_{ij}^{ab}}e^{-T}H_N e^{T}\kslat = 0,
            \label{eq:doubles_amplitude_1}
        \end{align}
        to find an expression that can be used to solve for the amplitudes 
        $t_{ij}^{ab}$.
        By employing the Baker-Campbell-Haussdorf (BCH) expansion,
        while setting $T= T_2$, we find
        \begin{equation}
                \bar{H}
                    = \Big(
                            H_N + H_NT_2
                            + \frac{1}{2}H_NT_2^2 \Big)_C.
        \label{eq:normal_order_expansion}
        \end{equation}
        The subscript $c$ indicates that only those terms in which the
        Hamiltonian is connected\footnote{Again, in aWick's theorem sense}
        to every cluster operator on its right should be included. Since
        the Hamiltonian contains at most four annihilation and creation
        operators, $H_N$ can connect to at most four cluster operators
        at once. Therefore, the BCH expansion must truncate at the
        fourth-order terms.

@@

        Then we can define a new version of \autoref{eq:CCD_amp_1},
        \begin{equation}
        \begin{aligned}
            g(u, t) =&
                    u^{ab}_{ij} + \tilde{f}^b_c t^{ac}_{ij}P(ab) - \tilde{f}^k_jt^{ab}_{ik}P(ij) \\
                 +& \frac{1}{4}t^{cd}_{ij} t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}t^{cd}_{ij} u^{ab}_{cd} \\
                 +& \frac{1}{2}t^{cd}_{jm} t^{ab}_{in} u^{mn}_{cd} P(ij) - \frac{1}{2}t^{ac}_{nm} t^{bd}_{ij} u^{nm}_{cd} P(ab) \\
                 +& t^{ac}_{im} t^{bd}_{jn} u^{mn}_{cd} P(ij) + t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
                 -& \frac{1}{2}t^{ab}_{im} u^{mn}_{jn}.
        \end{aligned}
        \end{equation}

@@

        \subsubsection{Intermediate Computations}
        Looking closely at the amplitude equation in (\ref{eq:CCD_amp_1}) one might come
        to realize that this equation contains many of the same structures in
        several of the terms. This warrants the search for an algebraic transformation of the
        CCD amplitude equation that has the potential to reduce the amount of floating point
        operations needed to compute it. As it turns out, such terms exist and they will decrease
        the computing time necessary by an order of magnitude. We will define the following
        "intermediates",
        \begin{align}
            \label{eq:intermediate1}
            \chi^{ab}_{cd} &= \frac{1}{4}t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}u^{ab}_{cd} \\
            \label{eq:intermediate2}
            \chi^n_j &= \frac{1}{2}t^{cd}_{jm} u^{mn}_{cd} \\
            \label{eq:intermediate3}
            \chi^a_d &= \frac{1}{2} t^{ac}_{nm} u^{nm}_{cd} \\
            \label{eq:intermediate4}
            \chi^{bm}_{jc} &= u^{bm}_{jc} + \frac{1}{2}t^{bd}_{jn}u^{mn}_{cd}
        \end{align}
        These intermediate structures will allow us to rewrite \autoref{eq:CCD_amp_1} to,
        \begin{equation}
        \begin{aligned}
            0 =& u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab) - f^k_jt^{ab}_{ik}P(ij) \\
              &+ t^{cd}_{ij}\chi^{ab}_{cd} + t^{ab}_{in}\chi^n_jP(ij)
               - t^{bd}_{ij}\chi^a_dP(ab) \\
              &+ t^{ac}_{im}\chi^{bm}_{jc}P(ab)P(ij) + \frac{1}{2}t^{ab}_{im}u^{mn}_{jn}.
        \end{aligned}
        \end{equation}
        The importance of this "trick" will become apparent in due time.

@@

        \subsubsection{Constructing the Hartree-Fock basis}
            Having found the matrix elements of $h$ and $u$ we can now use the
            \emph{self-consistent field} (SCF) iteration method to construct $h$ and
            $u$ in a \emph{restricted Hartree-Fock} (RHF) basis. This will yield a
            better estimate to the actual, unknown basis functions of the
            system.

            The neatest, yet arguably the most abstract, way to write the Hartree-Fock
            equations for electron $i$\footnote{This is weird, as electron
            should be indistinguishable and therefore impossible to label.} is
            \begin{equation}
                f_i \varphi_i = \varepsilon_i \varphi_i,
            \end{equation}
            where $f_i$ is the Fock operator, $\varphi_i$ are eigenstates of the Fock
            operator consiting of a set of one-electorn wave funcitons called the
            Hartree-Fock molecular orbitals, and $\varepsilon_i$ are the eigenenergies
            of the Fock operator. The fock operator, in matrix notation, is given by
            \begin{equation}
                F_{pa} = H_{pq} +  J(D)_{pq} - \frac{1}{2}K(D)_{pq},
            \end{equation}
            where $h_i$ is the one-body operatore, while the two-body operator is
            divided into what we call a direct part,
            \begin{equation}
                J(D)_{pq} = \braket{pq}{rs}D_{sr},
            \end{equation}
            and an exchange part,
            \begin{equation}
                K(D)_{pq} = \braket{ps}{rq}D_{sr}.
            \end{equation}
            The direct part of the two-body operator is comparable to classical Coloumb
            repulsion, while the exchange does not have a classical analog as it arises
            from the antisymmetry requirement of the wavefunction. $D_{sr}$ is the
            density matrix of the system.

@@

            Because the RHF SCF method is a variational method, the initial
            wavefunction made up of the single-particle wavefunctions $\varphi_i$
            is varied until the minimal energy is found. The transformations necessary
            to make all these variations will now be contained in $C$. So if we start from
            a harmonic oscillator (HO) basis we can transform into a better basis, called a
            Hartree-Fock (HF) basis,
            \begin{equation}
                \ket{\Psi}^{\text{HF}}_p = \sum_{k=1}^L C_{kp} \ket{\Psi}_k^{\text{HO}}.
            \end{equation}
            We will see how the coupled cluster method with double excitations works with
            both the HO basis and the HF basis, but not equally well.

@@

    Because we have employed a mixture of Python and C++ with Cython interfaces,
    the code sturcture may be difficult to undertand for someone without Cython
    experience, but we have tried to make the actual usage of the software as
    painless as possible.

    Most of the computations that are done in a coupled cluster algorithm
    is linear algebra and we found that NumPy, relying on
    BLAS and LAPACK, is most likely as fast as it gets and, at the very least,
    efficient enough for our purposes. The quicker and easier scripting capabilities
    Python provides is worth what little extra speed-up writing the entire
    code base in C++ could have provided.

@@

        The software is easy to install by first cloning the GitHub reopsitory,
        \begin{bash}
            git clone git@github.com:Schoyen/FYS4411.git
        \end{bash}
        Then all you need to do is change directory to the project code folder
        where a \lstinline{MakeFile} is included so you only need to build and
        install,
        \begin{bash}
            cd FYS4411/project_2/coupled-cluster
            make build
            make install
        \end{bash}
        Now everything should be able to run from anywhere on your computer.
        To ensure that all requirements of the program are satified you can
        install with \mintinline{bash}{make installr} instead.

        We have included a sample python script in \autoref{app:sample_script}
        that uses almost all methods and classes in the package, and also compares
        the time each methods takes to fininsh the computations.

@@

        Both in \mintinline{python}{CoupledClusterDoublesSparse} and
        \mintinline{python}{CoupledClusterDoublesOptimized} we make use
        of the intermediates in equations \ref{eq:intermediate1} to
        \ref{eq:intermediate4}. We have also relied upon implementations
        in C++ for some of the heaviest computations. In the optimized
        class special care was taken to ensure that every contraction
        was computed in an optimal way, with the minimization of memory
        use and CPU time as the overall goal. Moreover, several methods
        in NumPy for doing linear algebra was tested, including but not limited
        to \mintinline{python}{numpy.einsum()}, \mintinline{python}{numpy.tensordot},
        and reshaping datastructures to two dimensions and using 
        \mintinline{python}{numpy.dot}. Additionally, we tried to make use of 
        the python package \mintinline{python}{dask} which allows for arrays that
        are larger than the computer's memory buffer as well as automated
        concurrency. We found that the dask arrays was a bad fit for our code
        base, but we ended up using \mintinline{python}{numba} for automated
        paralellization of some iterative loops as well as just-in-time
        compialtion of some functions.

@@

        All the results from our simulations are displayed in tables \ref{tab:N2},
        \ref{tab:N6}, \ref{tab:N12} and \ref{tab:N20}. They show the ground state
        energy as a result of restricted Hartree-Fock SCF iterations, and coupled
        cluster doubles with harmonic oscillator basis and Hatree-Fock basis for 
        systems of increasing size, $N=2$, $N=6$, $N=12$ and $N=20$ particles
        respectively. The results are thoroughtly discussed in the next section.

@@

        \begin{table}
            \centering
            \caption{Here we look at $N = 12$ particles ($R = 3$ shells).  We did
            not achieve convergence using the harmonic oscillator basis for the
            lower frequency values and large number of shells. No convergence is
            marked as {\nan} in the table.}
            \begin{ruledtabular}
                \begin{tabular}{c|c|ccc}
                    $\omega$ & $R$ & RHF & CCD(HO) & CCD(HF) \\
                    \hline
                          & $3$ & $46.361130$ & $46.361130$ & $46.361130$ \\
                          & $4$ & $43.663267$ & $45.837079$ & $43.309845$ \\
                          & $5$ & $41.108851$ & $45.456883$ & $40.654710$ \\
                          & $6$ & $40.750512$ & \nan & $40.068340$ \\
                    $0.5$ & $7$ & $40.302719$ & \nan & $39.508500$ \\
                          & $8$ & $40.263752$ & \nan & $39.399128$ \\
                          & $9$ & $40.216688$ & \nan & $39.329311$ \\
                          & $10$ & $40.216252$ & \nan & $39.309409$ \\
                          & $11$ & $40.216195$ & \nan & $39.296007$ \\
                          & $12$ & $40.216165$ & \nan & $39.285968$ \\
                    \hline
                          & $3$ & $73.765549$ & $73.765549$ & $73.765549$ \\
                          & $4$ & $70.673849$ & $73.314476$ & $70.324250$ \\
                          & $5$ & $67.569930$ & $72.990679$ & $67.031096$ \\
                          & $6$ & $67.296869$ & \nan & $66.526677$ \\
                    $1.0$ & $7$ & $66.934745$ & \nan & $66.049564$ \\
                          & $8$ & $66.923094$ & \nan & $65.972157$ \\
                          & $9$ & $66.912244$ & \nan & $65.921205$ \\
                          & $10$ & $66.912035$ & \nan & $65.889281$ \\
                          & $11$ & $66.911365$ & \nan & $65.866715$ \\
                          & $12$ & $66.911364$ & \nan & $65.849776$ \\
                    \hline
                          & $3$ & $120.722260$ & $120.722260$ & $120.722260$ \\
                          & $4$ & $117.339642$ & $120.296556$ & $116.995036$ \\
                          & $5$ & $113.660396$ & $120.007146$ & $113.048934$ \\
                          & $6$ & $113.484866$ & $119.759037$ & $112.658821$ \\
                    $2.0$ & $7$ & $113.247601$ & $119.662199$ & $112.309482$ \\
                          & $8$ & $113.246579$ & $119.584733$ & $112.235521$ \\
                          & $9$ & $113.246303$ & $119.524394$ & $112.181828$ \\
                          & $10$ & $113.245854$ & $119.472283$ & $112.140661$ \\
                          & $11$ & $113.245256$ & $119.430353$ & $112.109973$ \\
                          & $12$ & $113.245183$ & $119.394712$ & $112.085683$ \\
                    \hline
                          & $3$ & $242.334879$ & $242.334879$ & $242.334879$ \\
                          & $4$ & $238.739591$ & $241.927593$ & $238.394598$ \\
                          & $5$ & $234.352741$ & $241.663950$ & $233.680649$ \\
                          & $6$ & $234.282331$ & $241.507595$ & $233.425243$ \\
                    $5.0$ & $7$ & $234.194820$ & $241.390525$ & $233.226529$ \\
                          & $8$ & $234.194059$ & $241.293417$ & $233.137933$ \\
                          & $9$ & $234.190797$ & $241.221056$ & $233.070205$ \\
                          & $10$ & $234.190714$ & $241.158896$ & $233.020009$ \\
                          & $11$ & $234.190665$ & $241.109926$ & $232.980634$ \\
                          & $12$ & $234.190553$ & $241.068091$ & $232.948528$
                \end{tabular}
            \end{ruledtabular}
            \label{tab:N12}
        \end{table}
